Embracing the Laws of Physics: \\ Reversible Models of Computation

### 2017-10-16 survey/outline.tex
lots and lots of comments.

\documentclass[12pt]{article}
\usepackage{fullpage}
\title{Embracing the Laws of Physics: \\ Reversible Models of Computation}
\author{Jacques Carette, Roshan James, Amr Sabry}
\begin{document}
\maketitle

@@

Cyber-physical systems increasingly common. Are computing foundations
adequate?

@@

@@

@@

@@

@@

@@

>>>>

\documentclass[12pt]{article}
\usepackage{fullpage}
\title{Embracing the Laws of Physics: \\ Reversible Models of Computation}
\author{Jacques Carette, Roshan James, Amr Sabry}

\newcommand{\amr}[1]{\fbox{Amr says:} \textbf{#1}}
\newcommand{\jc}[1]{\fbox{Jacques says:} \textbf{#1}}
\newcommand{\roshan}[1]{\fbox{Roshan says:} \textbf{#1}}

\begin{document}
\maketitle

@@

\jc{why are logic and DM listed alongside MoC?}

Cyber-physical systems increasingly common. Are computing foundations
adequate? 
\jc{your questioning of MoC below is not about CPS! Should either
foreshadow the next section 'properly' here, or add a leading 
example based on CPS to MoC section}

@@

\jc{While all the above are true (and not exhaustive), what I also read
is that the first two are actually ``the same'' point mostly related
to scale and distributivity, the 3rd is about inequate models (but not 
related to physics), while the 4th is about the physics of devices.
They are not about CPS. And not about Quantum either. So how are these
motivating a deeper look at how quantum affects our MoC? It feels to me
that we ought to have a quantum ``example'' here already.}

\jc{What this seems to motivate is to look at our models with respect
to what we're trying to achieve.  Models are, by definition, abstractions
that ``abstract away'' irrelevant details. So the point would be that
some details that we thought were irrelevant really are not.
So I could see this section motivating a whole programme of
targetted de-abstraction. This could lead into the focus of this paper
following one strand, namely one aspect of quantum physics.}

@@

\jc{No access to infinite memory/tape just means that ``infinitary''
arguments do not go through. The MoC with reversible computation 
assumes totally reliable operations (but not necessarily infinitely
many). You leap to ``continuum'' mathematics without justification.
While QC does use the continuum in non-trivial ways, and physical
reality is ``discrete'' at the atomic level (but who knows at the
quantum level?!?), I don't think that's where you want to go here.
But see below.}

@@

\jc{In a sense, physics already does this: statistical mechanics.
Even though the systems under study are explicitly discrete, albeit
will huge numbers of particules, the behaviour is approximated (!!)
via a continuous model. Same in economics with Black-Scholes.
The real problem are those conclusions drawn from such models which
*require* arbitrarily small values to exist, and even worse, for
the underlying values (reals) to be connected, i.e. form an actual
continuum. What Feynman is asking here is 'finiteness of information',
which is what things like Type-2 computability also ask for.}

@@

\jc{One must be careful with this argument: we have a ``current 
understanding'' of quantum physics, which could be quite flawed.
So what we want to do is to learn from thsoe parts of quantum physics
which seem to be extremely stable (like conservation of information!)
first. As researchers, we can certainly speculate on it all, but it
makes sense to vigorously pursue the conclusions one can draw from
those aspects of (quantum) physics that seem stable.}

@@

\jc{Should we also have a quick survey of other works in reversible
computation? Most of the other work I know tries to start from 
well known models of computation or well known programming languages,
and then adapts them to be reversible. Pi is different in that it 
starts from the semantic implications of reversibility. Then, because
it adds types, rather than control-flow (or even data-flow) as its 
next layer of ``understanding'', this leads to equivalences, isomorphisms,
etc.  This then leads, quite naturally, to finding that the ``proof
language'' of semirings (and Rig Groupoids at level 2) is actually a 
programming language. And it is Pi. This is a neat twist on Curry-Howard
because CH is about \textbf{inhabitation} only. But with ``conservation
of information'' as the basis, a different kind of correspondance
arises; in fact, this one may well be an actual isomorphism.}

@@

\jc{This is where I am still least sure. I am wary of either doing
new research with such a tight timeline, or of being too speculative.
If there are papers on this already, perhaps references to them could
be put here?}
